### 2017-11-08 authors.tex
changes

\author{Nick Merrill, Max T Curran, John Chuang}
\affiliation{%
  \institution{BioSENSE, UC Berkeley School of Information}
  \city{Berkeley, California} 
  \state{USA} 
}
\email{ffff, mtcurran, john.chuang@berkeley.edu}

\hypersetup{
 pdfauthor={},
 pdftitle={},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 25.1.1 (Org mode 9.0.4)}, 
 pdflang={English}}

\title{Is the Future of Authenticity In Our Heads?}
\subtitle{Moving Passthoughts From the Lab to the World}

\begin{abstract}
Passthoughts, in which a user thinks a secret thought to log in to services or devices, provides two factors of authentication (knowledge and inherence) in a single step. Since its proposal in 2005, passthoughts enjoyed a number of successful empirical studies. In this paper, we renew the promise of passthoughts authentication, outlining the main challenges that passthoughts must overcome in order to move from the lab to the real world. We propose two studies, which seek different angles at the fundamental questions we pose. Further, we propose it as a fruitful case study for thinking about what authentication can, and should, be expected to do, as it pushes up against questions of what sorts of ``selves'' authentication systems must be tasked with recognizing. Through this discussion, we raise novel possibilities for authentication broadly, such as ``organic passwords'' that change naturally over time, or systems that reject users who are not acting quite ``like themselves.''
\end{abstract}

%
% Here's a visual tool for generating the CCS categories LaTeX below:
% 
%    http://dl.acm.org/ccs/ccs.cfm
%
% Use it to generate the two b
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002978.10002991.10002992.10011619</concept_id>
<concept_desc>Security and privacy~Multi-factor authentication</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10002978.10003029.10003032</concept_id>
<concept_desc>Security and privacy~Social aspects of security and privacy</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Security and privacy~Multi-factor authentication}
\ccsdesc[300]{Security and privacy~Social aspects of security and privacy}

\keywords{passthoughts, authentication, usable security}